# Reproducible Research:  Assessment 1
## Loading and preprocessing the data
```{r loaddata}
fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
zipfile <- "./course5week2_data.zip"
fileDirectory <- "./Data"
unzip_path <- "./Data/unzip"  ##### path for storing the unzipped files #######
if (!file.exists(fileDirectory)){
  dir.create(fileDirectory)
}
download.file(fileurl,file.path(zipfile))
unzip(zipfile,exdir=unzip_path) ####### exdir is the extract directory ##########
datafile <- file.path(unzip_path,"activity.csv")
activity <- read.csv(datafile)
activity$weekend <- as.factor(ifelse(weekdays(as.Date(activity$date))=="Saturday" | weekdays(as.Date(activity$date))=="Sunday","weekend","weekday"))
activity$dayofweek <- as.factor(weekdays(as.Date(activity$date)))
```

## What is mean total number of steps taken per day?
### Histogram of the total number of steps taken per day
```{r}
stepsByDay <- activity %>% group_by(date) %>% summarise(stepsperday = sum(steps,na.rm = TRUE))
qplot(stepsperday,data=stepsByDay,na.rm=TRUE,binwidth=500,xlab='Total steps per day', ylab='Frequency using binwith 500',main = 'Histogram of the total number of steps taken each day')
```

### Mean and median of the total number of steps taken per day
````{r}
meanstepsperday <- stepsByDay %>% summarise(average = mean(stepsperday,na.rm = TRUE),median=median(stepsperday,na.rm = TRUE))
meanstepsperday

```

## What is the average daily activity pattern?
### Time series plot of the 5-minute interval and the average number of steps across all days
```{r}
interval_average <- activity %>% group_by(interval) %>% summarise(average = mean(steps,na.rm = TRUE))
qplot(interval,average,data=interval_average,geom="line",xlab = "5-minute intervals",ylab = "Average steps taken across all days")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
interval_average[which.max(interval_average$average),]

```

## Imputing missing values
### Number of missing values in the dataset
```{r}
activity_na <- activity[which(is.na(activity$steps)),]
nrow(activity_na)
```

### Fill NAs with average steps based on interval
```{r}
# subset dataset where there are no NAs
    activity_no_NA <- activity[which(!is.na(activity$steps)),]

# calculate the mean steps for each interval
interval_only <- activity_no_NA %>% group_by(interval) %>% summarise(average=mean(steps))

# convert the average to integer
interval_only$average <- as.integer(interval_only$average)

activity_na$steps <- ifelse(activity_na$interval==interval_only$interval,interval_only$average)

```

### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activity_impute <- rbind(activity_no_NA,activity_na)
```

### Histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.
```{r}
#Histogram of the total number of steps taken each day after missing values are imputed
stepsByDay_impute <- activity_impute %>% group_by(date) %>% summarise(stepsperday = sum(steps))
qplot(stepsperday,data=stepsByDay_impute,na.rm=TRUE,binwidth=500,xlab='Total steps per day', ylab='Frequency using binwith 500',main = 'Histogram of the total number of steps taken each day')

#Mean and median number of steps taken each day
totalstepsperday_impute <- activity_impute %>% group_by(date) %>% summarise(stepsperday = sum(steps))
mean_n_median <- totalstepsperday_impute %>% summarise(average=mean(stepsperday),median=median(stepsperday))
mean_n_median

```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
meansteps <- activity_impute %>% group_by(interval,weekend) %>%   summarise(average = mean(steps))
qplot(interval,average,data=meansteps,geom="line",facets=weekend~.,xlab="5-minute interval",ylab="average number of steps",main="Average steps pattern between Weekday and Weekend")
```


